Workflow for Perth-Peel
========================================================

This document describes the steps that are needed to get from the raw data to a zonation run.  Note that all files should currently be pointing at data at a resolution of 100m.

```{r source function}
setwd("~/RSP/R Code/Perth-Peel")

ksource <- function(x, ...) {
  library(knitr)
  source(purl(x, output = tempfile()), ...)
}
```

1. Extract data from the downloaded files from NatureMap and the Atlas of Living Australia

```{r extract_data}
ksource("PP_extract records from ALA and NM.rmd")

```

2. Create environmental layers
a. Create a clipping mask using all spatial areas to be included
b. Generate percentage cover layers from nvis
c. Create individual rasters for each TEC type by attributes
d. Resample all environmental layers to match

```{r resample environment}
ksource("create clipping mask for zonation runs.Rmd")

ksource("PP_percentage cover.Rmd")

ksource("PP_extract TEC raster by attributes.Rmd")

ksource("PP_resample rasters to make identical.Rmd")

```

3.  Create input species files for Maxent and Zonation
- Species with greater than the minimum number of records exported to maxent.data.csv and maps to pdf
- Tiffs created for species of special interest
- TECs listed and added to zonation spp files


```{r create Maxent files}

ksource("PP_create maxent and zonation files.Rmd")

# use this if you need to reclip the files for some reason but don't want to start from the raw data (i.e. you changed the veg mask)
ksource("PP_convert ssi txt files to tif rasters.Rmd")

```

4. Run Maxent from batch file
- Run synctoy batch file to keep data on SC synchronised

5. Convert Maxent output ascii files to tifs and clip by extant vegetation
- Note that there are dates in this file to point to the correct maxent folder - will need to change these to make sure that we end up with the right data at the end.

6. Create pdfs of Maxent output and point data for DEC and SEWPaC

```{r clip maxent files}

ksource("PP_ascii files to tifs.Rmd")

ksource("PP_save Maxent sdms to pdf.rmd")

```

6. Run zonation models from batch file
- Run synctoy batch file to keep data on SC synchronised

7. Create development scenarios
- write a script that takes the shapefiles/rasters of potential development scenarios, merges them together and masks them to the SA region
```{r create development scenarios}
