Create extant vegetation layer to clip all layers for zonation
========================================================

Create a clippling mask that is used to cut the MaxEnt output for use in Zonation.  Needs to include all areas that we are interested in prioritising too and currently includes the following layers:
* remnant vegetation (DEC)
* TECs
* EPA bio3
* EPA bio2
* pines
* plantations(?)
* Ramsar
* world, commonwealth & national heritage


```{r setup_functions}
#load packages and set working directory
rm(list=ls())
start.time <- proc.time()

envi.directory <- "~/GIS_data/Perth-Peel/"
output.directory <- "~/GIS_data/Perth-Peel/environmental data/100m/scenarios/"

setwd(output.directory)

packages(raster)
packages(rgdal)
packages(maptools)

source("C:/Users/awhitehead/Dropbox/Random R code/shp2raster.R")

  GDA94 <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
  GDA94.50 <- CRS("+proj=utm +zone=50 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")

perth <- raster("~/GIS_data/Perth-Peel/environmental data/100m/region.mask.asc")
  perth[!is.na(perth)] <- 0
  writeRaster(perth,"~/GIS_data/Perth-Peel/mask files/perth.mask.tif", format="GTiff", overwrite=T)

PP.mask <- readShapePoly("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/DOP/DoP_StrategicAssessmentStudyBoundary_Extended_20121004_GDA94.shp")

create.raster <- function (s, mask.raster, label, value, transform=TRUE) {
    
   if(transform==TRUE) {
     proj4string(s) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
    s <- spTransform(s, CRS("+proj=utm +zone=50 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))
     }
    
  r <- rasterize(s,mask.raster)
    r[!is.na(r)] <- value
    r <- mask(merge(r,mask.raster),mask.raster)
    names(r) <- label
    
    plot(r, main=label)
    
   writeRaster(r,paste0(output.directory,label), format="GTiff", overwrite=T)
    return(r)
  }

```

1.  Create an empty raster the same resolution, extent and projection as the region
2.  Load in desired layers and set background cells to NA.
3.  Check to make sure that the extents match and resample if they don't.
4.  Stack layers together and merge, setting all non-NA cells to 1.
5.  Export clipping mask for future analyses.

```{r read in spatial data that is to be included in mask}
#vegetation
nvis_mvs_exotics <- raster("~/GIS_data/Perth-Peel/environmental data/100m/ibra/vegetation/nvis_mvs_exotics.tif")
  nvis_mvs_exotics[nvis_mvs_exotics == 98 | nvis_mvs_exotics == 99] <- NA
epa.bio3 <- raster("EPA_Biodiversity3_20130322.tif")
  epa.bio3[epa.bio3 == 0] <- NA
rem.veg <- raster("DEC_RemnantVegetation_SwanRegion_20120401.tif")
  rem.veg[rem.veg == 0] <- NA
# pines <- raster("~/GIS_data/Perth-Peel/environmental data/100m/ibra/vegetation/pines.tif")
  #pines[!is.na(pines)] <-1

# Heritage
world.heritage <- raster("PP_WHA_clip.tif")
  world.heritage[world.heritage == 0] <- NA
national.heritage <- raster("PP_NHL_clip.tif")
  national.heritage[national.heritage == 0] <- NA
commonwealth.heritage <- raster("PP_CHL_clip.tif")
  commonwealth.heritage[commonwealth.heritage == 0] <- NA
  
# Wetlands
ramsar <- raster("PP_Ramsar_clip.tif")
  ramsar[ramsar == 0] <- NA
epa.bio2 <- raster("EPA_Biodiversity2_20130322.tif")
  epa.bio2[epa.bio2 == 0] <- NA

# migratory bird areas
mig.birds <- readShapePoly("~/GIS_data/Perth-Peel/DEC/DEC_SPB_ThreatenedandMigratoryBirds_20130308.shp")
  mig.birds.raster <- create.raster(mig.birds, perth,value=1, label="DEC_SPB_ThreatenedandMigratoryBirds_20130308", transform=T )
  mig.birds.raster[mig.birds.raster == 0] <- NA

# TECs
tec <- raster("~/GIS_data/Perth-Peel/environmental data/100m/tecs/tecs_100m")
cec <- raster("~/GIS_data/Perth-Peel/SEWPaC/PP_EC_data/PP_EC")
saltmarsh <- raster("~/GIS_data/Perth-Peel/zonation/100m/EC_saltmarsh.tif")


```

```{r read in spatial data that is to be excluded from mask}
existing.development <- readShapePoly("~/GIS_data/Perth-Peel/DOP/DoP_Development_Footprint_20140310.shp")
  exist.dev.raster <- shp2raster(existing.development[grepl("Developed|Existing Development",existing.development$DEVELOPED),], perth,value=0,bkg=1, label="DoP_Development_Footprint_20140310_developed", transform=F )

existing.infrastructure <- readShapePoly("~/GIS_data/Perth-Peel/DOP/DoP_Infrastructure_Existing_Footprint_20140321.shp",proj4string=CRS("+proj=tmerc +lat_0=0 +lon_0=117 +k=0.9996 +x_0=500000 +y_0=4000000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))
  exist.infra.raster <- shp2raster(existing.infrastructure, perth, value=0, bkg=1, label="DoP_Infrastructure_Existing_Footprint_20140321", transform=T, proj.to=GDA94.50)

```


```{r merge layers}
rasters <- list(epa.bio3,world.heritage,national.heritage,commonwealth.heritage,ramsar,epa.bio2,rem.veg,tec,cec,saltmarsh,mig.birds.raster)

# check to make sure that the extents match
for (i in 1:length(rasters)){
  r <- rasters[[i]]
 cat("Checking",names(r),"\n")
 if(extent(r)!= extent(perth)){
   cat("Resampling",names(r),"\n")
   r <- resample(r,perth, method="ngb",filename=names(r),overwrite=T)
  assign(names(r),r)
   }
}

all.include <- extend(calc(stack(ramsar,tec,cec,saltmarsh,mig.birds.raster,epa.bio3,world.heritage,national.heritage,commonwealth.heritage,epa.bio2,rem.veg),max),nvis_mvs_exotics)
origin(nvis_mvs_exotics) <- origin(all.include)
all.include.nvis <- mosaic(all.include,nvis_mvs_exotics,fun=max)
  all.include.nvis[!is.na(all.include.nvis)] <- 1
  

# combine all inclusion layers
all.include <- mask(calc(stack(ramsar,tec,cec,saltmarsh,mig.birds.raster,epa.bio3,world.heritage,national.heritage,commonwealth.heritage,epa.bio2,rem.veg,nvis_mvs_exotics),max),perth)
  all.include[!is.na(all.include)] <- 1

# combine all exclusion layers
all.exclude <- min(exist.dev.raster,exist.infra.raster)

# combine inclusion & exclusion layers
clipping.mask <- min(all.include,all.exclude)
  clipping.mask[clipping.mask==0] <- NA

plot(clipping.mask)

#writeRaster(clipping.mask,"~/GIS_data/Perth-Peel/mask files/perth.clipping.mask.tif", format="GTiff", overwrite=T)

# export clipping mask that does not exclude existing development as these still contain areas of native vegetation
writeRaster(all.include,"~/GIS_data/Perth-Peel/mask files/perth.clipping.mask.tif", format="GTiff", overwrite=T)
writeRaster(all.include.nvis,"~/GIS_data/Perth-Peel/mask files/ibra.clipping.mask.tif", format="GTiff", overwrite=T)
```


*This file was last updated on 25 June 2014 and last run on `r format(Sys.time(), "%d %B %Y")`.*