Create development rasters for scenario planning
========================================================

This script takes the relevant shapefiles from the various planning agencies in Perth-Peel and creates mask rasters for each of the 12 scenarios tested in the draft report using the function `create.raster`.  Shapefiles are forced into the perth boundary at 100m and then clipped by the clipping layer to exclude non-vegetated areas.

Raster values range from 1-5 to force the order in which zonation removes cells
1 = areas for development
2 = general areas
3 = areas with the lowest form of protection
4 = areas that are sort of protected (ie. state forests)
5 = areas that have a high level of protection


```{r setup}
#load packages and set working directory
rm(list=ls())
start.time <- proc.time()


  envi.directory <- "Z:/Amy/GIS_data/Perth-Peel/zonation/environmental data/100m/scenario data/"
  output.directory <- "Z:/Amy/GIS_data/Perth-Peel/zonation/100m/scenarios/"

setwd(output.directory)

library(raster)
library(rgdal)
library(maptools)
library(rgeos)

perth <- raster("Z:/Amy/GIS_data/Perth-Peel/environmental data/100m/region.mask.asc")
  perth[!is.na(perth)] <- 2
clipping.mask <- raster("Z:/Amy/GIS_data/Perth-Peel/environmental data/100m/clipping.mask.tif")
  clipping.mask[!is.na(clipping.mask)] <- 2

```

The variables in the create.raster function are defined below:
s = input shapefile
mask.raster = the background raster that you are converting into
raster.label = label that is used as the filename, raster name and plot title
value = the value that cells are to be given where the shapefile is present in the landscape
transform = a logical value that determines whether the function transforms the shapefile data prior to conversion.  If set to TRUE, then you need to make sure that the CRS values are set correctly.

```{r create raster function}
create.raster <- function (s, mask.raster, raster.label, value, transform=TRUE) {
    
   if(transform==TRUE) {
     proj4string(s) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
    s <- spTransform(s, CRS("+proj=utm +zone=50 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))
     }
    
  r <- rasterize(s,mask.raster)
    r[!is.na(r)] <- value
    r <- mask(merge(r,mask.raster),mask.raster, filename=paste0(output.directory,raster.label), format="GTiff", overwrite=T)
    names(r) <- rep(raster.label,length(names(r)))
    
    plot(r, main=raster.label, zlim=c(1,5))

    return(r)
  }

```


```{r DOP EIA1}

dop.eia1 <- readShapePoly(paste0(envi.directory,"DOP/DoP_EIA1_20130809.shp"))

dop.eia1.proposed <- create.raster(dop.eia1[dop.eia1$Land_Type!="Developed Urban" & dop.eia1$Land_Type!="Existing Industrial",],perth, raster.label="DOP_EIA1_proposed", value=1, transform=FALSE)

```

```{r BRM EIA1}

dmp.eia1 <- readShapePoly(paste0(envi.directory,"DMP/DMP_BRM_EIA_V1.shp"))

dmp.eia1 <- create.raster(dmp.eia1, perth, raster.label="DMP_EIA1", value=1, transform=FALSE)

```

```{r pines}
pines.shp <-readShapePoly(paste0(envi.directory,"DEC/DEC_FPC_Pines_gnangara_pinjar_and_yanchep-standing_2011-12.shp"))

pines <- create.raster(pines.shp, perth, raster.label="Pines_gnangara_pinjar_and_yanchep", value=1, transform=FALSE)

```

```{r DOP EIA2}
dop.eia2 <- readShapePoly(paste0(envi.directory,"DOP/DoP_EIA2_20130809.shp"))

dop.eia2.proposed <- create.raster(dop.eia2[dop.eia2$Land_Type!="Developed Urban" & dop.eia2$Land_Type!="Existing Industrial",], perth, raster.label="DOP_EIA2_proposed", value=1, transform=FALSE)

```

This isn't included in the scenarios
```{r BRM EIA2}

dmp.eia2 <- readShapePoly(paste0(envi.directory,"DMP/DMP_BRM_EIA_V2.shp"))

dmp.eia2 <- create.raster(dmp.eia2, perth, raster.label="DMP_EIA2", value=1, transform=FALSE)

```

```{r BRM EIL V4}
dmp.eil4 <- readShapePoly(paste0(envi.directory,"DMP/DMP_BRM_EIL_V4.shp"))

dmp.eil4.raster <- create.raster(dmp.eil4, perth, raster.label="DMP_EILV4",value=1,transform=FALSE) 

```

EIA3 has two options but needs to be combined with the Industrial Lands Strategy to be a fair comparison with EIA1 & EIA2 (which include industrial and urban)

```{r DOP EIA3 + industrial}

# Option 1 without Keralup
dop.EIA3_Option1_30102013 <- readShapePoly(paste0(envi.directory,"DOP/DOP_EIA3_Option1_30102013.shp"))

dop.EIA3_Option1_30102013.raster <- create.raster(dop.EIA3_Option1_30102013, perth, raster.label="DOP_EIA3_Option1",value=1,transform=FALSE)
  

# Option 2 with Keralup
dop.EIA3_Option2_30102013 <- readShapePoly(paste0(envi.directory,"DOP/DOP_EIA3_Option2_30102013.shp"))

dop.EIA3_Option2_30102013.raster <- create.raster(dop.EIA3_Option2_30102013, perth, raster.label="DOP_EIA3_Option2",value=1,transform=FALSE)

# Industrial Land Strategy
dop.industrial <- readShapePoly(paste0(envi.directory,"DOP/Industrial_Data/IndustrialLandStrategy_08Nov2013.shp"))

dop.industrial.raster <- create.raster(dop.industrial, perth, raster.label="DOP_IndustrialLandStrategy",value=1,transform=FALSE)

# merge industrial land strategy and EIA3 Options 1 & 2
eia3_ils_Option1 <- calc(stack(dop.EIA3_Option1_30102013.raster,dop.industrial.raster),fun=min, filename=paste0(output.directory,"eia3_ils_Option1"), format="GTiff", overwrite=T)
  plot(eia3_ils_Option1)

eia3_ils_Option2 <- calc(stack(dop.EIA3_Option2_30102013.raster,dop.industrial.raster),fun=min, filename=paste0(output.directory,"eia3_ils_Option2"), format="GTiff", overwrite=T)
  plot(eia3_ils_Option2)

```

The infrastructure layers are based on a list provided by Leo Peter.  Note that points need to be buffered by a small amount to create polygons that can be used in `create.raster`.

```{r infrastructure}

dot.proposed.railstops <- readShapePoints(paste0(envi.directory,"DOT/DoT_PassengerRailStops_Proposed_20121017.shp"))
  dot.proposed.railstops <- gBuffer(dot.proposed.railstops,width=10, byid=TRUE)
future.railstops <- create.raster(dot.proposed.railstops, perth, raster.label="DoT_PassengerRailStops_Proposed",value=1,transform=FALSE)

Future_Roads_Metropolitan_Peel_DoP <- readShapeLines(paste0(envi.directory,"DOP/Future_Roads_Metropolitan_Peel_DoP.shp"))
  #Future_Roads_Metropolitan_Peel_DoP <- gBuffer(Future_Roads_Metropolitan_Peel_DoP,width=100,byid=TRUE)
future.roads <- create.raster(Future_Roads_Metropolitan_Peel_DoP, perth, raster.label="Future_Roads_Metropolitan_Peel_DoP",value=1,transform=FALSE)

WP_New_Sub_stations_20131003 <- readShapePoints(paste0(envi.directory,"WP/WP_New_Sub_stations_20131003.shp"))
  WP_New_Sub_stations_20131003 <- gBuffer(WP_New_Sub_stations_20131003,width=10, byid=TRUE)
  future.substations <- create.raster(WP_New_Sub_stations_20131003, perth, raster.label="WP_New_Sub_stations",value=1, transform=FALSE)

WP_New_Power_Lines_20131003 <- readShapeLines(paste0(envi.directory, "WP/WP_New_Power_Lines_20131003.shp"))
  #WP_New_Power_Lines_20131003 <- gBuffer(WP_New_Power_Lines_20131003, width=100, byid=TRUE)
  future.powerlines <- create.raster(WP_New_Power_Lines_20131003, perth, raster.label="WP_New_Power_Lines_20131003", value=1, transform=FALSE)

WC_future_assests_20131010 <- readShapeLines(paste0(envi.directory,"WC/WC_future_assests_20131010.shp"))
  #WC_future_assests_20131010 <- gBuffer(WC_future_assests_20131010, width=100, byid=TRUE)
future.water <- create.raster(WC_future_assests_20131010, perth, raster.label = "WC_future_assests_20131010", value= 2, transform=FALSE)

DoT_RecreationalPaths_Future_20121127 <- readShapeLines(paste0(envi.directory,"DOT/DoT_RecreationalPaths_Future_20121127.shp"))
  #DoT_RecreationalPaths_Future_20121127 <- gBuffer(DoT_RecreationalPaths_Future_20121127, width=100, byid=TRUE)
future.paths <- create.raster(DoT_RecreationalPaths_Future_20121127, perth, raster.label="DoT_RecreationalPaths_Future_20121127", value=1, transform=FALSE)

DoT_PrincipalSharedPaths_Future_20121127 <- readShapeLines(paste0(envi.directory,"DOT/DoT_PrincipalSharedPaths_Future_20121127.shp"))
shared.paths <- create.raster(DoT_PrincipalSharedPaths_Future_20121127,perth,raster.label="DoT_PrincipalSharedPaths_Future_20121127",value=1,transform=FALSE)

DoT_PassengerRailExtension_Proposed_20121017 <- readShapeLines(paste0(envi.directory,"DOT/DoT_PassengerRailExtension_Proposed_20121017.shp"))
  #DoT_PassengerRailExtension_Proposed_20121017 <- gBuffer(DoT_PassengerRailExtension_Proposed_20121017, width=100, byid=TRUE)
future.rail <- create.raster(DoT_PassengerRailExtension_Proposed_20121017, perth, raster.label="DoT_PassengerRailExtension_Proposed_20121017", value=1, transform=FALSE)

SMPSP_road_network_conceptual_20131108 <- readShapeLines(paste0(envi.directory,"SMPSP/SMPSP_road_network_conceptual_20131108.shp"),delete_null_obj=TRUE)
  #SMPSP_road_network_conceptual_20131108 <- gBuffer(SMPSP_road_network_conceptual_20131108, width=100, byid=TRUE)
future.smpsp.roads <- create.raster(SMPSP_road_network_conceptual_20131108, perth, raster.label="SMPSP_road_network_conceptual_20131108", value=1, transform=FALSE)

future.infrastructure <- calc(stack(future.railstops, future.roads, future.substations, future.powerlines, future.water, future.paths, shared.paths,future.rail,future.smpsp.roads),fun=min, filename=paste0(output.directory,"infrastructure"), format="GTiff", overwrite=T)
  plot(future.infrastructure, main="future.infrastructure",zlim=c(1,3))

```


```{r rural residential}
rural_residential_zoned <- readShapePoly(paste0(envi.directory,"DOP/rural_residential_zoned.shp"))

rural.residential <- create.raster(rural_residential_zoned,perth, raster.label="rural_residential_zoned",value=1, transform=FALSE)

```


```{r create_conservation_rasters}
    
# DPAW Regional Parks under Section 16 of CALM Act
  regional.parks16 <- readShapePoly(paste0(envi.directory, "DEC/DEC_RegionalParks_withS16_20110502.shp"))
  regional.parks5 <- create.raster(regional.parks16,perth, raster.label="DEC_RegionalParks_withS16_20110502", value=5)

# DPAW Managed Lands
managed.lands <- readShapePoly(paste0(envi.directory, "DEC/DEC_ManagedLandsAndWaters_20120601.shp"))
    managed.lands$IUCN[managed.lands$CATEGORY=="State Forest"] <- 6
    
    managed.lands3 <- create.raster(managed.lands[is.na(managed.lands$IUCN),],perth, raster.label="DEC_ManagedLandsAndWaters_20120601_No IUCN status", value=3)
    
    managed.lands <- managed.lands[!is.na(managed.lands$IUCN),]
    
    # State Forests (IUCN 6)
    managed.lands4 <- create.raster(managed.lands[managed.lands$IUCN=="6",],perth, raster.label="DEC_ManagedLandsAndWaters_20120601_IUCN 6", value=4)
    
  # IUCN 1-4
    managed.lands5 <- create.raster(managed.lands[(managed.lands$IUCN=="1a" | managed.lands$IUCN=="2" | managed.lands$IUCN=="3" | managed.lands$IUCN=="4"),],perth, raster.label="DEC_ManagedLandsAndWaters_20120601_IUCN 1-4", value=5)

# Heritage
world.heritage <- readShapePoly(paste0(envi.directory, "SEWPaC/PP_MNES_extract/PP_WHA_clip.shp"))
  world.heritage5 <- create.raster(world.heritage,perth, raster.label="PP_WHA_clip", value=5)

national.heritage <- readShapePoly(paste0(envi.directory, "SEWPaC/PP_MNES_extract/PP_NHL_clip.shp"))
  national.heritage5 <- create.raster(national.heritage,perth, raster.label="PP_NHL_clip", value=5)

commonwealth.heritage <- readShapePoly(paste0(envi.directory, "SEWPaC/PP_MNES_extract/PP_CHL_clip.shp"))
  commonwealth.heritage5 <- create.raster(commonwealth.heritage,perth, raster.label="PP_CHL_clip", value=5)

# Bush Forever
bushforever.SP28 <- readShapePoly(paste0(envi.directory,"DOP/DoP_StatePlanningPolicy28_BushForeverReserves_20120823.shp"))
  bushforever3 <- create.raster(bushforever.SP28,perth, raster.label="DoP_StatePlanningPolicy28_BushForeverReserves_20120823", value=3, transform=FALSE)

#MRS
parks.rec <- readShapePoly(paste0(envi.directory,"MRS_PRS/mrspoly_20120928.shp"))
  parks.rec3 <- create.raster(parks.rec[(parks.rec$DESCRIPTIO=="PARKS & RECREATION" | parks.rec$DESCRIPTIO=="PARKS & RECREATION RESTRICTED"), ],perth, raster.label="mrspoly_20120928", value=3, transform=FALSE)

#PRS
open.space <- readShapePoly(paste0(envi.directory,"MRS_PRS/prspoly_20120928.shp"))
  open.space3 <- create.raster(open.space[open.space$DESCRIPTIO=="REGIONAL OPEN SPACE" , ],perth, raster.label="prspoly_20120928", value=3, transform=FALSE)

protected.areas <- calc(stack(regional.parks5,managed.lands3,managed.lands4,managed.lands5,world.heritage5,national.heritage5,commonwealth.heritage5,bushforever3,parks.rec3,open.space3),fun=max, filename=paste0(envi.directory,"protected.areas"), format="GTiff",overwrite=T)
  plot(protected.areas, main="Protected areas", zlim=c(1,5))

```

Create the merged scenario files and save with the final scenario names.

```{r scenarios}

scenario2_eia1 <-  writeRaster(dop.eia1.proposed,filename=paste0(output.directory,"scenario2_eia1"),format="GTiff",overwrite=T)
  plot(scenario2_eia1, main="scenario2_eia1", zlim=c(1,5))

scenario3_eia2 <-  writeRaster(dop.eia2.proposed,filename=paste0(output.directory,"scenario3_eia2"),format="GTiff",overwrite=T)
  plot(scenario3_eia2, main="scenario3_eia2", zlim=c(1,5))

scenario4_eia3_Option1 <-  writeRaster(eia3_ils_Option1,filename=paste0(output.directory,"scenario4_eia3_Option1"),format="GTiff",overwrite=T)
  plot(scenario4_eia3_Option1, main="scenario4_eia3_Option1", zlim=c(1,5))

scenario5_eia3_Option2 <-  writeRaster(eia3_ils_Option2,filename=paste0(output.directory,"scenario5_eia3_Option2"),format="GTiff",overwrite=T)
 plot(scenario5_eia3_Option2, main="scenario5_eia3_Option2", zlim=c(1,5))

scenario6_brm_eil4 <- writeRaster(dmp.eil4.raster,filename=paste0(output.directory,"scenario6_brm_eil4"),format="GTiff",overwrite=T)
  plot(scenario6_brm_eil4, main="scenario6_brm_eil4", zlim=c(1,5))

scenario7_rural_res <- writeRaster(rural.residential,filename=paste0(output.directory,"scenario7_rural_res"),format="GTiff",overwrite=T)
  plot(scenario7_rural_res, main="scenario7_rural_res", zlim=c(1,5))

scenario8_infrastructure <- writeRaster(future.infrastructure,filename=paste0(output.directory,"scenario8_infrastructure"),format="GTiff",overwrite=T)
  plot(scenario8_infrastructure, main="scenario8_infrastructure", zlim=c(1,5))

scenario9_pines <- writeRaster(pines,filename=paste0(output.directory,"scenario9_pines"),format="GTiff",overwrite=T)
  plot(scenario9_pines, main="scenario9_pines", zlim=c(1,5))

scenario10_protected.areas <- writeRaster(protected.areas,filename=paste0(output.directory,"scenario10_ProtectedAreas"),format="GTiff",overwrite=T)
  plot(scenario10_protected.areas, main="scenario10_protected.areas", zlim=c(1,5))

scenario11_All.Option1 <- calc(stack(eia3_ils_Option1,dmp.eil4.raster,rural.residential,future.infrastructure,pines),min,filename=paste0(output.directory,"scenario11_AllOption1"),format="GTiff",overwrite=T)
  plot(scenario11_All.Option1, main="scenario11_All.Option1", zlim=c(1,5))

scenario12_All.Option2 <- calc(stack(eia3_ils_Option2,dmp.eil4.raster,rural.residential,future.infrastructure,pines),min,filename=paste0(output.directory,"scenario12_AllOption2"),format="GTiff",overwrite=T)
  plot(scenario12_All.Option2, main="scenario12_All.Option2", zlim=c(1,5))


```

All the scenarios need to be clipped by the vegetation layer to remove the areas that are already developed and, thus, not being considered.


```{r clip scenarios by clipping mask}

(scenarios <- dir(output.directory,pattern="scenario.*\\.tif$"))

for (s in seq(scenarios)){
  r <- raster(scenarios[s])
  r <- mask(r,clipping.mask,filename=paste0("Z:/Amy/GIS_data/Perth-Peel/zonation/100m/",names(r)), format="GTiff", overwrite=T)
  plot(r, main=names(r),zlim=c(1,5))
}

```



*This file was last updated on 22 November 2013 and last run on `r format(Sys.time(), "%d %B %Y")`.*