Create data files for Maxent and Zonation analyses
========================================================

This script takes the downloaded data from the [Australian Living Atlas](http://www.ala.org.au/species-by-location/) and [NM Atlas](http://www.environment.NM.gov.au/atlaspublicapp/UI_Modules/ATLAS_/AtlasSearch.aspx) for all specified species, as well as the data from the brut dataset, and produces:   
1. a pdf of maps of all species locations (if `save.maps == TRUE`)   
2. a csv file of location points for input into Maxent `maxent.data`   
3. a spp file for input into Zonation that lists all of the species ascii files and other attributes `zonation.spp`  

The first part of the script takes the brut dataset where species have been coded into taxonomic groups and identifies species that have more than the specified minimum number of records that fall within the Hunter region as masked by a shapefile. We extract all records for these species from within the NM North Coast and Sydney Basin bioregions  

Then we run a loop that runs through the species listed as *Vulnerable* or *Endangered* under the NM legislation on the NM database that have more than the minumim number of records and extracts the latitude and longitude of all records in the ALA and NM databases.   

Data are only retained if they occur after a specified date.


```{r setup, fig.width=7, fig.height=6, message=FALSE, warning=FALSE, comment="", tidy=TRUE}
#load packages and set working directory
rm(list=ls())
# source("C:/Users/awhitehead/Dropbox/Random R code/tweet me.r")

packages(ggplot2)
packages(ggmap)
packages(maptools)
packages(sp)
packages(rgdal)
packages(raster)

setwd("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/species data/")

# set output options
  save.maps <- FALSE  # save maps as pdf
  save.tifs <- FALSE # save SSI files as tifs
  save.txt <- TRUE # save SSI files as text files
  save.output <-TRUE # save maxent data as csv file
  include.brut <- FALSE # include brut dataset in maxent file

combine.records <- function (NC.records, SB.records) {
  tryCatch({
            rbind(NC.records,SB.records)
            }, error = function(err){
              if(is.null(NC.records)==TRUE & is.null(SB.records)==FALSE){
                return(SB.records)
              } else if(is.null(SB.records)==TRUE & is.null(NC.records)==FALSE){
                return(NC.records)
              } else cat("")
            })
  }

# set the google map background for the species plots and create background map
#   map <- get_map(location = c(115.85,-32.18), zoom = 7, maptype="terrain")
  PP.mask <- readShapePoly("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/DOP/DoP_StrategicAssessmentStudyBoundary_Extended_20121004_GDA94.shp")
    PP.mask.data <- fortify(PP.mask)

  ibra.mask <- readShapePoly("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/environmental data/PP_IBRA_gda94.shp")
    ibra.mask.data <- fortify(ibra.mask)
    
#   print(base.map <- ggmap(map) + geom_polygon(aes(x = long, y = lat, group = group), data=PP.mask.data, colour = 'yellow', fill = 'black', alpha = .1, size = .3) + geom_polygon(aes(x = long, y = lat, group = group), data=ibra.mask.data, colour = 'orange', fill = 'black', alpha = .1, size = .3))  

# The minimum number of records required for a species to be included  
  min.records <- 20
  
# The earliest date for selecting records
  min.date <- strptime("01/01/1990", "%d/%m/%Y")

# Minimum allowable spatial accuracy value (combined number of characters in lat & long ~100m)
  min.accuracy <- 14

# open species list
  protected.species <- read.csv("~/GIS_data/Perth-Peel/species data/WA listed protected species.csv")

# remove species that are supposed to be extinct & species to ignore
threatened.species <- protected.species[protected.species$ignore != "Y",]

#identify which species have more than the minimum number of records across both atlases and are not listed as species to ignore (i.e. marine species)
species.list <- threatened.species$Scientific.Name[rowSums(threatened.species[,c("ALA.records","NM.records","DEC.records")],na.rm=TRUE) >0]

# add additional DEC species
# dec.species <- c("Eucalyptus argutifolia","Grevillea curviloba subsp. curviloba", "Grevillea curviloba subsp. incurva", "Grevillea flexuosa", "Lasiopetalum pterocarpum", "Lepidosperma rostratum", "Macarthuria keigheryi", "Synaphea sp. Fairbridge Farm (D. Papenfus 696)", "Synaphea sp. Pinjarra (R. Davis 6578)", "Thelymitra dedmaniarum", "Verticordia fimbrilepis subsp. fimbrilepis")

# create empty dataframes for the output data
  maxent.data <- d(species=NULL,longitude=NULL,latitude=NULL, database=NULL)
  species.data.list <- NA
  ssi.list <- NA

```

Species are kept for analysis in Maxent if they have more than `r min.records` records within the Hunter region that were observed after `r min.date`, with a minimum accuracy of ~ 100m. 

```{r brut_analysis, message=FALSE, warning=FALSE, comment="", tidy=TRUE}

if(include.brut == TRUE) {

    setwd("C:/Users/awhitehead/Documents/GIS_data/Hunter/species point data/")
  
  # open the brut dataset and generate spatial co-ordinates
    modeled.sp <- read.csv("C:/Users/awhitehead/Documents/GIS_data/Australia-wide/species_data/modeledsp_order_brut.csv")
      modeled.sp <- modeled.sp[!is.na(modeled.sp$YEARmax >= (min.date$year+1900)),]
      modeled.sp$database <- "brut"
    coordinates(modeled.sp) <- c("Lon_Centre","Lat_Centre")
  
    brut.maxent.data <- maxent.data
  
  # clip the brut data by the region mask, provided there are at least 30 points in the hunter region
    ibra.sp <- modeled.sp[ibra.mask,]
    hccrem.sp <- modeled.sp[hccrem.mask,]
  
  # identify how many observations for each species within the hunter
   n.obs <- d(obs=tapply(hccrem.sp$SEL,hccrem.sp$TAXON_ID,sum))
      n.obs$taxa <- rownames(n.obs)
      n.obs <- n.obs[n.obs$obs > min.records,]
  
  # extract each species with more than the minimum number of records from brut dataset 
    for (i in seq(n.obs$taxa)){
        input <- as.data.frame(hccrem.sp[hccrem.sp$TAXON_ID==n.obs$taxa[i],])
      colnames(input)[3:4] <- c("latitude", "longitude")
      input$species <- paste(input$Ordertot,input$TAXON_ID)
      
  #     if(nrow(input) > min.records){
        brut.maxent.data <- rbind(brut.maxent.data,input[,c("species","longitude", "latitude", "databse")])
        sp.map <- base.map + geom_point(aes(x = longitude, y = latitude,color=database,shape=species), data = input, alpha=0.5)
        print(sp.map)
        cat("\n", "Adding", nrow(input), "records for", as.character(input$species[i]), "to maxent data")
  #     }
    }
    
}

```

```{r ALA_NM_analysis, message=FALSE, warning=FALSE, comment="", tidy=TRUE}
setwd("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/species data/")
atlas.maxent.data <- maxent.data
ssi.zonation.data <- maxent.data

# set specifications for pdf file if save.maps == TRUE
    if(save.maps == TRUE) pdf("species point data maps.pdf",paper="a4r")

# run loop to extract data from ALA and NM atlas data by species 
for (i in seq(species.list)){
  ALA.NM.data <- NULL
  
  # extract ALA data
  if (!is.na(threatened.species$Downloaded.ALA[threatened.species$Scientific.Name == species.list[i]]) == TRUE){
    ALA.data <- read.csv(paste0("ALA data/",species.list[i],".csv"))
      
      ALA.labels <- c("Latitude...processed", "Longitude...processed")
    colnames(ALA.data)[which(names(ALA.data) %in% ALA.labels)] <- c("latitude", "longitude")
    ALA.data$index <- paste(sprintf("%.6f",ALA.data$latitude), sprintf("%.6f",ALA.data$longitude))
    ALA.data$database <- "ALA"
    ALA.data$species <- species.list[i]
    ALA.data$accuracy.digits <- nchar(ALA.data$latitude,type="chars") + nchar(ALA.data$longitude,type="chars")
    ALA.data <- ALA.data[ALA.data$accuracy.digits >=min.accuracy,]
    ALA.data <- ALA.data[!is.na(ALA.data$Year...parsed >= (min.date$year+1900)),]
  } else ALA.data <- NULL

  #extract NM Atlas data
  if (!is.na(threatened.species$Downloaded.NM[threatened.species$Scientific.Name == species.list[i]]) == TRUE){
    NM.data <- read.csv(paste0("NatureMap data/",species.list[i],".csv"))
    NM.data$database <- "NM"
    NM.data$species <- species.list[i]
    NM.labels <- c("LONGITUDE", "LATITUDE")
    colnames(NM.data)[which(names(NM.data) %in% NM.labels)] <- c("longitude","latitude")
    NM.data$index <- paste(sprintf("%.6f",NM.data$latitude), sprintf("%.6f",NM.data$longitude))
    NM.data <- NM.data[NM.data$COLDATEY >= (min.date$year + 1900) & !is.na(NM.data$COLDATEY),]
    NM.data$accuracy.digits <- nchar(NM.data$latitude,type="chars") + nchar(NM.data$longitude,type="chars")
    NM.data <- NM.data[NM.data$accuracy.digits >=min.accuracy,]
  } else NM.data <- NULL
  
  ALA.NM.data <- combine.records(ALA.data[,c("species","longitude","latitude","database")],NM.data[,c("species","longitude","latitude","database")])
  
  
# add CC roosting sites  
  if(species.list[i] == "Calyptorhynchus latirostris"){
    CC.data <- read.csv("DEC_Carnabys_roost_sites.csv")
      CC.data <- CC.data[CC.data$site_categ!="CLEARED",]
    CC.data$species <- "Calyptorhynchus latirostris"
    CC.data$database <- "DEC"
    CC.labels <- c("decimal_la", "decimal_lo")
    colnames(CC.data)[which(names(CC.data) %in% CC.labels)] <- c("latitude","longitude")
    
   ALA.NM.data <- combine.records(ALA.NM.data[,c("species","longitude","latitude","database")],CC.data[,c("species","longitude","latitude","database")]) 
      
  }
  
  if (!is.na(threatened.species$Downloaded.DEC[threatened.species$Scientific.Name == species.list[i]]) == TRUE){
    dec.data <- read.csv(paste0("DEC data/",species.list[i],".csv"))
      colnames(dec.data[2]) <- "species"
    dec.data$database <- "DEC"
    dec.labels <- c("TAXON1", "Latitude", "Longitude")
    colnames(dec.data)[which(names(dec.data) %in% dec.labels)] <- c("species","latitude","longitude")
    protected.species$final.records[protected.species$Scientific.Name == species.list[i]] <- nrow(dec.data)
          
     ALA.NM.data <- combine.records(ALA.NM.data[,c("species","longitude","latitude","database")],dec.data[,c("species","longitude","latitude","database")])
     } 
  
  
  if(nrow(ALA.NM.data) > 0) {
    species.data.list <- append(species.data.list, species.list[i])
    coordinates(ALA.NM.data) <- c("longitude", "latitude")
    
    ibra.sp <- ALA.NM.data[ibra.mask,]
      ibra.sp <- as.data.frame(ibra.sp)
      in.ibra <- nrow(ibra.sp)
    
    # only include species if more than the minimum number of records occur in the SA area
     PP.sp <- ALA.NM.data[PP.mask,]
      PP.sp <- as.data.frame(PP.sp)  
      in.PP <- nrow(PP.sp)
    
    # produce maps of the distribution of each species and save data
    # if more than minimum records save to maxent data
    # if less than minimum records but not zero, save to species of special interest
    if(in.PP >= min.records) {
#       try({
#         sp.map <- base.map + geom_point(aes(x = longitude, y = latitude,color=database, shape=species), data = ibra.sp, alpha=0.5)
#       print(sp.map)
#         }, silent=TRUE)
    
    # output data for maxent analyses
      atlas.maxent.data <- rbind(atlas.maxent.data,ibra.sp)
      cat("\n", "Adding", in.ibra, "records for", as.character(species.list[i]), "to maxent data")
  
      } else {
        if(0 < in.PP & in.PP < min.records) {
          # list species with less than minimum records for inclusion as species of special interest
          ssi.list <- append(ssi.list,as.character(unique(PP.sp$species)))
          ssi.zonation.data <- rbind(ssi.zonation.data, ibra.sp)
          cat("\n", "Adding", in.ibra, "records for", as.character(species.list[i]), "as species of special interest")
        }
        }
    
    # remove all species-specific datafiles, so they don't mess up the next round of the loop
    suppressWarnings(rm(ALA.data, NM.data, dec.data, ALA.NM.data, PP.sp,in.PP,sp.map, ibra.sp,in.ibra))
    gc()
  }
}

# close pdf file
  if(save.maps == TRUE) dev.off()

```


These are combined together in `maxent.data` where there are `r nrow(maxent.data)` records for `r (length(species.list) + n.species)` species.  Note that there may be duplicate records in these files but they should be filtered out by Maxent. A map of the distribution of each species is saved in a pdf file (*C:/Users/awhitehead/Documents/GIS_data/Hunter/species point data/species point data maps.pdf*)

```{r export_maxent.data, message=FALSE, warning=FALSE, comment="", tidy=TRUE}
if(include.brut == TRUE) {
    gda_maxent.data <- rbind(brut.maxent.data[,c("species","longitude","latitude")], atlas.maxent.data[,c("species","longitude","latitude")])
    } else gda_maxent.data <- atlas.maxent.data[,c("species","longitude","latitude")]

h(gda_maxent.data)

# convert maxent data to GDA94 Transverse Mercator Zone 56 to match the rest of the data
  coordinates(gda_maxent.data) <- c("longitude", "latitude")
  
  # set current projection
  proj4string(gda_maxent.data) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

  # transform to new projection
  project_maxent.data <- spTransform(gda_maxent.data, CRS("+proj=utm +zone=50 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))

  # convert back to a dataframe
  gda.tm_maxent.data <- as.data.frame(project_maxent.data)
    names(gda.tm_maxent.data)[2:3] <- c("easting","northing")

  h(gda.tm_maxent.data)

  # remove duplicate records for each species (i.e. keep only one for each grid cell)
  ibra.raster <- raster("C:/Users/awhitehead/Documents/GIS_Data/Perth-Peel/environmental data/pp_ibra_mask.tif")

  export_maxent.data <- maxent.data
  export_ssi.zonation.data <- maxent.data # catch the species that get dropped due to duplicate removal
  
    for(i in seq(unique(gda.tm_maxent.data$species))){
      input <- gda.tm_maxent.data[gda.tm_maxent.data$species==unique(gda.tm_maxent.data$species)[i],]
      cell <- cellFromXY(ibra.raster,input[,2:3])
      output <- input[!duplicated(cell),]
      if(nrow(output)>min.records){
        export_maxent.data <- rbind(export_maxent.data,output)
        protected.species$final.records[protected.species$Scientific.Name==unique(gda.tm_maxent.data$species)[i]] <- nrow(output)
        protected.species$sdm[protected.species$Scientific.Name==unique(gda.tm_maxent.data$species)[i]] <- "TRUE"
        } else export_ssi.zonation.data <- rbind(export_ssi.zonation.data,output)
      protected.species$final.records[protected.species$Scientific.Name==unique(gda.tm_maxent.data$species)[i]] <- nrow(output) 
      rm(input,cell,output)
    }

export.species <- unique(export_maxent.data$species)
                                  
if(save.output ==TRUE) {
  write.csv(export_maxent.data, "C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/Maxent_files/species_data/maxent.data.csv",row.names=FALSE)
  
  taxa <- unique(protected.species$Taxa)
  
  for(t in seq(taxa)){
    taxa.species <- intersect(protected.species$Scientific.Name[which(protected.species$Taxa==taxa[t])],export_maxent.data$species)
    taxa.maxent <- export_maxent.data[export_maxent.data$species %in% taxa.species,]
    write.csv(taxa.maxent,paste0("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/Maxent_files/species_data/maxent.data_",taxa[t],".csv"),row.names=FALSE )
    rm(taxa.species, taxa.maxent)
  }
  
  write.csv(protected.species, "C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/species data/WA listed protected species.csv", row.names=FALSE)
  

cat("\n", nrow(export_maxent.data), "records for", length(export.species), "species exported to maxent.data.csv from pool of",length(species.data.list), "species with available data")
  }
```

```{r ssi_data, message=FALSE, warning=FALSE, comment="", tidy=TRUE}
pp.clipping.mask <- raster("~/GIS_data/Perth-Peel/environmental data/100m/clipping.mask.tif")

if(include.brut == TRUE) {
    gda_ssi.zonation.data <- rbind(brut.ssi.zonation.data[,c("species","longitude","latitude")], atlas.ssi.zonation.data[,c("species","longitude","latitude")])
    } else gda_ssi.zonation.data <- ssi.zonation.data[,c("x","y","species")]

h(gda_ssi.zonation.data)

# convert maxent data to GDA94 Transverse Mercator Zone 50 to match the rest of the data
  coordinates(gda_ssi.zonation.data) <- c("x", "y")
  
  # set current projection
  proj4string(gda_ssi.zonation.data) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

  # transform to new projection
  project_ssi.zonation.data <- spTransform(gda_ssi.zonation.data, CRS("+proj=utm +zone=50 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))

  # convert back to a dataframe
  gda.tm_ssi.zonation.data <- as.data.frame(project_ssi.zonation.data)
    names(gda.tm_ssi.zonation.data)[1:2] <- c("easting","northing")

  h(gda.tm_ssi.zonation.data)
  
    for(i in seq(unique(gda.tm_ssi.zonation.data$species))){
      input <- gda.tm_ssi.zonation.data[gda.tm_ssi.zonation.data$species==unique(gda.tm_ssi.zonation.data$species)[i],]
      cell <- cellFromXY(clipping.mask,input[,2:3])
      output <- input[!duplicated(cell),]
      export_ssi.zonation.data <- rbind(export_ssi.zonation.data,output)
        protected.species$final.records[protected.species$Scientific.Name==unique(gda.tm_ssi.zonation.data$species)[i]] <- nrow(output) 
      rm(input,cell,output)
    }

export_ssi.zonation.data$biological.value <- 1.0
export_ssi.zonation.data$uncertainty <- 0

ssi.export.species <- unique(export_ssi.zonation.data$species)

                                  
# save data for species with greater than the minimum number of records after the minimum date in the Hunter region to one file for Maxent analysis
if(save.txt ==TRUE) {
    for (i in seq(ssi.export.species)) {
      if(ssi.export.species[i]!="Pseudemydura umbrina"){
      export <- export_ssi.zonation.data[export_ssi.zonation.data$species == ssi.export.species[i],]
      write.table(export[,!names(export) %in% c("species")], paste0("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/zonation/100m/ssi_files/",ssi.export.species[i],".txt"),row.names=FALSE, sep="\t", col.names=FALSE)
      }
    }
  }
    
if(save.tifs ==TRUE) {
    presence.absence.raster <- function (mask.raster,species.data,raster.label="") {
    #set the cells that contain points to 1
  speciesRaster <- rasterize(species.data,mask.raster,field=1)
    speciesRaster <- mask(merge(speciesRaster,mask.raster),mask.raster)
    #label the raster
    names(speciesRaster) <- raster.label
   writeRaster(speciesRaster,paste0("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/zonation/100m/",raster.label,".tif"), format="GTiff",overwrite=T)
  cat("Converted",raster.label,"to raster","\n")
}
    
   for (i in seq(ssi.export.species)) {
     if(ssi.export.species[i]!="Pseudemydura umbrina"){
       export <- export_ssi.zonation.data[export_ssi.zonation.data$species == ssi.export.species[i],]
        export <- export[,2:3]
        species <- as.character(gsub(" ","_",ssi.export.species[i]))
        presence.absence.raster(clipping.mask,export, raster.label=species)
     }
    } 
       
cat("\n", nrow(export_ssi.zonation.data), "records exported to", length(ssi.export.species), "individual txt & tif files for ssi from pool of",length(species.data.list), "species with available data")
  }

write.csv(protected.species, "C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/species data/WA listed protected species.csv", row.names=FALSE)

```

```{r tec_data}
#extract details of tecs for zonation species files
 community.list <- threatened.species$Scientific.Name[threatened.species$ignore=="N" ]

# append tec names to ssi list
ssi.export.species <- append(as.character(ssi.export.species), as.character(community.list))

```

The `zonation.spp` table lists each species and the required input for zonation.  These are currently all set to `1` and will require some tinkering as we get further down the line. This assumes that we will run all the listed species in Maxent from `maxent.data.csv`, creating all the relevant ascii files. *Or we will use this same data in biomod2.*

```{r export_zonation.file, message=FALSE, warning=FALSE, comment="", tidy=TRUE}
  
  zonation.spp <- d(weight=rep(1.0,length(export.species)), alpha=1.0, bqp.row=1.0, bqp.buffer=1.0, cell.removal=1.0, species=NA )
    zonation.spp$species <- paste0(export.species,".tif")
    zonation.spp$species <- gsub(" ","_",zonation.spp$species)

mnes.export.species <- export.species[which(export.species %in% protected.species$Scientific.Name[protected.species$mnes == TRUE])]

mnes.zonation.spp <- d(weight=rep(1.0,length(mnes.export.species)), alpha=1.0, bqp.row=1.0, bqp.buffer=1.0, cell.removal=1.0, species=NA )
    mnes.zonation.spp$species <- paste0(mnes.export.species,".tif")
    mnes.zonation.spp$species <- gsub(" ","_",mnes.zonation.spp$species)

h(zonation.spp)
  
# species of special interest
ssi.zonation.spp <- d(weight=rep(1.0,length(ssi.export.species)), alpha=1.0, bqp.row=1.0, bqp.buffer=1.0, cell.removal=1.0, species=NA )
  ssi.zonation.spp$species <- paste0(ssi.export.species,".tif")
  ssi.zonation.spp$species <- as.character(gsub(" ","_",ssi.zonation.spp$species))

# mnes species of special interest
mnes.ssi.list <- ssi.export.species[which(ssi.export.species %in% protected.species$Scientific.Name[protected.species$mnes == TRUE])]
mnes.ssi.zonation.spp <- d(weight=rep(1.0,length(mnes.ssi.list)), alpha=1.0, bqp.row=1.0, bqp.buffer=1.0, cell.removal=1.0, species=NA )
  mnes.ssi.zonation.spp$species <- paste0(mnes.ssi.list,".tif")
  mnes.ssi.zonation.spp$species <- as.character(gsub(" ","_",mnes.ssi.zonation.spp$species))

all.zonation.spp <- rbind(zonation.spp,ssi.zonation.spp)
all.species <- all.zonation.spp$species
  
# weight by DoE weighting scheme
weighted.zonation.spp <- all.zonation.spp
  weighted.species <- as.character(gsub("_"," ",all.zonation.spp$species))
  weighted.species <- as.character(gsub(".tif","",weighted.species,fixed=TRUE))

  for(w in seq(weighted.species)){
    weighted.zonation.spp$weight[w] <- protected.species$rank_DoE[protected.species$Scientific.Name==weighted.species[w]]
    }

power <- c(1,1.5,2,3,4,5)
#change relative weighting of ranks
for (i in seq(power)){
  w <- weighted.zonation.spp
  w$weight <- w$weight^power[i]
  assign(paste0("rank",power[i],".zonation.spp"),w)
  write.table(w,paste0("~/GIS_data/Perth-Peel/zonation/100m/rank",power[i],".zonation.spp"),sep="\t",row.names=FALSE, col.names=FALSE)
  rm(w)
}

# weight by status only
weighted.status.zonation.spp <- all.zonation.spp
  weighted.status.species <- as.character(gsub("_"," ",all.zonation.spp$species))
  weighted.status.species <- as.character(gsub(".tif","",weighted.status.species,fixed=TRUE))

  for(w in seq(weighted.status.species)){
    weighted.status.zonation.spp$weight[w] <- protected.species$weight_status[protected.species$Scientific.Name==weighted.status.species[w]]
    }

#change relative weighting of ranks
for (i in seq(power)){
  w <- weighted.status.zonation.spp
  w$weight <- w$weight^power[i]
  assign(paste0("threat",power[i],".zonation.spp"),w)
  write.table(w,paste0("~/GIS_data/Perth-Peel/zonation/100m/threat",power[i],".zonation.spp"),sep="\t",row.names=FALSE, col.names=FALSE)
  rm(w)
} 

if(save.output ==TRUE) {
  write.table(zonation.spp,paste0("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/zonation/100m/zonation.spp"),sep="\t",row.names=FALSE, col.names=FALSE)
    write.table(mnes.zonation.spp,paste0("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/zonation/100m/mnes.zonation.spp"),sep="\t",row.names=FALSE, col.names=FALSE)
  write.table(ssi.zonation.spp,paste0("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/zonation/100m/ssi.zonation.spp"),sep="\t",row.names=FALSE, col.names=FALSE)
  write.table(mnes.ssi.zonation.spp,paste0("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/zonation/100m/mnes.ssi.zonation.spp"),sep="\t",row.names=FALSE, col.names=FALSE)
  write.table(all.zonation.spp,paste0("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/zonation/100m/all.zonation.spp"),sep="\t",row.names=FALSE, col.names=FALSE)
}

cat("\n", "zonation.spp file saved with default values for all",length(export.species),"species", "\n", "mnes.zonation.spp file saved with default values for all",length(mnes.export.species),"species","\n", "ssi.zonation.spp file saved with default values for all",length(ssi.list),"species","\n", "mnes.ssi.zonation.spp file saved with default values for all",length(mnes.ssi.list),"species")
```


```{r group_files}
groups.file <- d(species=all.species, output.group=1,condition=-1,retention.group=-1,retention.mode=-1,local.edge.correction=-1)

# MNES species == Group 2
groups.file$output.group[which(groups.file$species %in% mnes.ssi.zonation.spp$species)] <- 2

write.table(groups.file[,-1], paste0("C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/zonation/100m/listed_groups.txt"),sep="\t",row.names=FALSE, col.names=FALSE)


# write summary of run to log file
cat("\n","Date:",format(Sys.time(), " %d%m%Y"),"\n", "Total species:",nrow(all.zonation.spp),"\n","Minimum records:", min.records,"\n", "Maxent species:", nrow(zonation.spp), "\n","MNES Maxent species:",nrow(mnes.zonation.spp),"\n","SSI species:",nrow(ssi.zonation.spp)-length(community.list),"\n","MNES SSI species:", nrow(mnes.ssi.zonation.spp),"\n","SSI species exported as txt & tif files","\n", "TEC:", length(community.list),"\n", file="C:/Users/awhitehead/Documents/GIS_data/Perth-Peel/zonation/100m/LOG FILE.txt", append=T )

dmSend("I'm done creating the Maxent files","@nzwormgirl")

```
*This file was last updated on 15 July 2013 and last run on `r format(Sys.time(), "%d %B %Y")`.*
