My attempt at Neighbourhood Analysis in R
========================================================

I want to calculate the percentage of cells that contain a particular type of habitat within a given radius based on the Greater Hunter Vegetation Map.  This is my attempt at doing that.

The plan is to:  
1. create presence/absence rasters for each of the vegetation groups we are interested in  
2. calculate the sum of cells within a given radius  
3. recalculate this as a percentage  

The original data will need to be exported from the geodatabase as shapefiles and/or rasters before analysis.  

The following functions set up the circular filter window `make_circ_filter`, turn a raster in to presence-absence based on a threshold `presence.absence` and calculate the percentage cover within a given window `percent.cover`.  `make_circ_filter` uses code from [Michael Scroggie](http://scrogster.wordpress.com/2012/10/05/applying-a-circular-moving-window-filter-to-raster-data-in-r/) and requires a cell resolution and radius that is a multiple of the resolution. 


```{r chunkFunctions, warning=FALSE,message=FALSE}
rm(list=ls())
source("C:/Users/awhitehead/Dropbox/Random R code/tweet me.r")

packages(raster)
packages(rgdal)
packages(maptools)

#function to make a circular weights matrix of given radius and resolution
#NB radius must be an even multiple of res!
make_circ_filter<-function(radius, res){
  require(raster)
  circ_filter<-matrix(NA, nrow=1+(2*radius/res), ncol=1+(2*radius/res))
  dimnames(circ_filter)[[1]]<-seq(-radius, radius, by=res)
  dimnames(circ_filter)[[2]]<-seq(-radius, radius, by=res)
  sweeper<-function(mat){
    for(row in 1:nrow(mat)){
      for(col in 1:ncol(mat)){
        dist<-sqrt((as.numeric(dimnames(mat)[[1]])[row])^2 +
          (as.numeric(dimnames(mat)[[1]])[col])^2)
        if(dist<=radius) {mat[row, col]<-1}
      }
    }
    return(mat)
  }
out<-sweeper(circ_filter)
return(out)
}

# function to sum the values in the circular filter and divides by the number of cells 
##(removing all cells with NA values)
percent.cover <- function(x){ sum(x,na.rm=T)/sum(cf, na.rm=T)*100}

# function to convert a raster to presence-absence data based on a threshold
  # set threshold to zero to create a mask of the area
presence.absence <- function(x) { x[x<threshold] <- 0
                     x[x>=threshold] <- 1
                     return(x) }

```

```{r Analysis_function, message=FALSE}

# open the selected sample file
  generate.pc.ascii <- function (filename, mask) {
#     sample.raster <- raster(filename)
      sample.raster[!is.na(sample.raster)] <- 1
    
    cat("\n","Loaded",names(sample.raster))
    
#     sample.raster <- resample(sample.raster,mask, method="bilinear")

  # calculate the percentage cover within the circular filter
    cat("\n","Calculating % cover of",names(sample.raster)," within",radius,"m radius")
    r_filt <- mask(focal(sample.raster, w=cf, fun=percent.cover),mask)
  
    par(mfrow=c(1,2))
    plot(sample.raster, main=names(sample.raster), axes=F) #original data
    plot(r_filt, main=paste0(radius,"m ",names(sample.raster)), axes=F) #filtered data
    return(r_filt)  
  
  }
```

```{r create_raster}

create.raster <- function (s, mask.raster, label, value, transform=TRUE) {
    
   if(transform==TRUE) {
     proj4string(s) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
    s <- spTransform(s, CRS("+proj=utm +zone=50 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))
     }
    
  r <- rasterize(s,mask.raster)
    r[!is.na(r)] <- value
   if(extent(r) != extent(mask.raster)){ 
    r <- mask(merge(resample(r,mask.raster),mask.raster),mask.raster)
   } else r <- mask(merge(r,mask.raster),mask.raster)
    names(r) <- label
    
    plot(r, main=label)
    
   #writeRaster(r,paste0(output.directory,label), format="GTiff", overwrite=T)
    return(r)
  }

# not sure what this projection string is for
# "+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=0 +lon_0=134 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
```


```{r create nvis rasters}
  setwd("~/GIS_data/Perth-Peel/environmental data")

# nvis_mvs <- raster("~/GIS_data/Perth-Peel/environmental data/100m/vegetation/nvis_mvs_buff")
nvis_mvs <- raster("~/GIS_data/Perth-Peel/environmental data/100m/ibra/vegetation/ibra_nvis_mvs")
  nvis_mvs.names <- as.data.frame(nvis_mvs@data@attributes)
    nvis_mvs.names <- rbind(nvis_mvs.names, c(100,NA,"Plantations"))
    nvis_mvs.names <- rbind(nvis_mvs.names, c(101,NA,"Pines"))                          
nvis_mvs_classes <- read.csv("~/GIS_data/Perth-Peel/environmental data/100m/ibra/vegetation/nvis_mvs_classes.csv")
  nvis_mvs.names$classes <- nvis_mvs_classes$mvs_name[which(nvis_mvs_classes$mvs_num %in% nvis_mvs.names$ID)]
  nvis_mvs.names$woody <- nvis_mvs_classes$woody[which(nvis_mvs_classes$mvs_num %in% nvis_mvs.names$ID)]

nvis.mask <- nvis_mvs
# Banksia woodland based on DEC EC nomination
dec.banksia <- readShapePoly("~/GIS_Data/Perth-Peel/DEC/SCP_Rem_Wdl_DSM02.shp")
  banksia.raster <- create.raster(dec.banksia, nvis_mvs, label="Banksia_woodland", value=50, transform=TRUE)

nvis_mvs <- banksia.raster

  

# exotic plantations
plantations <- raster("~/GIS_data/Perth-Peel/environmental data/100m/ibra/vegetation/plantations.tif")

# plantations <- readShapePoly("~/GIS_data/Perth-Peel/DEC/Plantation/State/plantation_annual_report.shp")
# 
#   plantations <- spTransform(plantations,CRS("+proj=utm +zone=50 +south +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))

pines<-raster("~/GIS_data/Perth-Peel/environmental data/100m/ibra/vegetation/pines")


exotic <- c("plantations", "pines")

for (i in seq(exotic)) {
r <- get(exotic[i])
  if (extent(r)!=extent(nvis_mvs)) {
    cat("\n","Resampling",names(r))
    r <- mask(resample(r,nvis_mvs),nvis_mvs)
#     writeRaster(r,paste0("~/GIS_data/Perth-Peel/environmental data/100m/ibra/vegetation/",names(r),".tif"),overwrite=T)
    assign(names(r),r)
    }
}

plantations[!is.na(plantations)] <- 100
pines[!is.na(pines)] <- 101
  all.veg <- stack(nvis_mvs,pines)
    nvis_mvs_exotics <- round(max(all.veg,na.rm=T),0)
  writeRaster(nvis_mvs_exotics, "~/GIS_data/Perth-Peel/environmental data/100m/ibra/vegetation/nvis_mvs_exotics.tif", overwrite=T)

veg_types <- c("Eucalyptus","Melaleuca","Other shrublands","grassland","bare","Casuarina","Banksia","Cleared","Pines")

for(i in seq(veg_types[1:8])){
  classes <- grep(veg_types[i],nvis_mvs.names$classes)
  IsBecomes <- d(is=nvis_mvs.names$ID,becomes=NA)
    IsBecomes$becomes[classes] <- 1
    
  r <- reclassify(nvis_mvs,IsBecomes)
  names(r) <- veg_types[i]
  plot(r, main=veg_types[i])
  
  if(veg_types[i] == "Other shrublands") {
    assign("Other_shrublands",r)
    } else assign(veg_types[i],r)
  cat("\n","Creating",as.character(veg_types[i]), "raster")
}

# characterise woody vegetation
cat("\n","Creating Woody raster")
classes <- which(nvis_mvs.names$woody=="yes")
  
  IsBecomes <- d(is=nvis_mvs.names$ID,becomes=NA)
    IsBecomes$becomes[classes] <- 1
    
  Woody <- mask(reclassify(nvis_mvs_exotics,IsBecomes),nvis_mvs)
  names(Woody) <- "Woody"
  plot(Woody, main="Woody")
  

veg_types <- c("Eucalyptus","Melaleuca","Other_shrublands","grassland","bare","Casuarina","Banksia","Cleared","Woody","pines")

```

```{r calculate percent_cover}
#nvis_mvs_exotics <- raster("~/GIS_data/Perth-Peel/environmental data/100m/ibra/vegetation/nvis_mvs_exotics.tif")
        
# get the resolution of the mask raster
  resolution <- xres(nvis_mvs)
  
# ## Generate ascii files with percentage cover within 1000 m radius
#   radius <- resolution*10
# 
# # set the circular filter
#   cf<-make_circ_filter(radius, resolution)
# 
# 
# for(i in seq(veg_types)) {  
#   sample.raster <- get(veg_types[i])
#   output <- generate.pc.ascii(sample.raster,nvis_mvs)
#   writeRaster(output, paste0("100m/ibra/vegetation/",veg_types[i],radius,".tif"), overwrite=TRUE) 
#   rm(output)
#  }

## Generate ascii files with percentage cover within 2000 m radius
radius <- resolution*20

# set the circular filter
  cf<-make_circ_filter(radius, resolution)

for(i in seq(veg_types)) {  
  sample.raster <- get(veg_types[i])
  output <- generate.pc.ascii(sample.raster,nvis_mvs)
  writeRaster(output, paste0("100m/ibra/vegetation/",veg_types[i],radius,".tif"), overwrite=TRUE) 
  rm(output)
 }

```

```{r CBC}

# Carnaby's feeding habitat - Pines and Banksia (should this include anything else?)
cbc.food.JF <- readShapePoly("~/GIS_data/Perth-Peel/DEC/DEC_Carnabys_FeedAreasRequiringInvestigation_JF_IBRA_20101028.shp")
cbc.food.SCP <- readShapePoly("~/GIS_data/Perth-Peel/DEC/DEC_Carnabys_FeedAreasRequiringInvestigation_SCP_IBRA_20101028.shp")

cbc.food.JF.raster <- create.raster(cbc.food.JF, PP_mask,value=1,label="DEC_Carnabys_FeedAreasRequiringInvestigation_JF_IBRA_20101028", transform=F)
cbc.food.SCP.raster <- create.raster(cbc.food.SCP, PP_mask,value=1,label="DEC_Carnabys_FeedAreasRequiringInvestigation_JF_IBRA_20101028", transform=F)
cbc.food.raster <- calc(stack(cbc.food.JF.raster, cbc.food.SCP.raster),sum)
writeRaster(cbc.food.raster, "~/GIS_data/Perth-Peel/environmental data/100m/vegetation/DEC_Carnabys_FeedAreasRequiringInvestigation_all.tif", overwrite=T )



plot(cbc.food.raster)

nvis_mvs.pines <- raster("~/GIS_data/Perth-Peel/environmental data/100m/vegetation/nvis_mvs.pines.asc")
cbc.feeding <- nvis_mvs.pines
  cbc.feeding[cbc.feeding!=100 & cbc.feeding!=50] <- 0
  cbc.feeding[cbc.feeding>0] <-1
writeRaster(cbc.feeding, "~/GIS_data/Perth-Peel/environmental data/100m/vegetation/cbc.feeding.BP.tif", overwrite=T )



dmSend("I'm done creating the percentage cover rasters", "@nzwormgirl")

```


*This file was last updated on 08 July 2013 and last run on `r format(Sys.time(), "%d %B %Y")`.*