Create development rasters for scenario planning
========================================================

This script takes the relevant shapefiles from the various planning agencies in Perth-Peel and creates mask rasters for each of the 12 scenarios tested in the draft report using the function `create.raster`.  Shapefiles are forced into the perth boundary at 100m and then clipped by the clipping layer to exclude non-vegetated areas.

Raster values range from 1-5 to force the order in which zonation removes cells
1 = areas for development
2 = general areas
3 = areas with the lowest form of protection
4 = areas that are sort of protected (ie. state forests)
5 = areas that have a high level of protection


```{r setup}
#load packages and set working directory
rm(list=ls())
start.time <- proc.time()

  computer <- "E:/Super Computer Data/"
  # computer <- "Z:/Amy/"

  envi.directory <- paste0(computer, "GIS_data/Perth-Peel/")
  output.directory <- paste0(computer, "GIS_data/Perth-Peel/zonation/100m/Conservation_Letters/final/scenarios/")

setwd(output.directory)

library(raster)
library(rgdal)
library(maptools)
library(rgeos)

perth <- raster(paste0(computer,"GIS_data/Perth-Peel/mask files/perth.mask.tif"))
  perth[!is.na(perth)] <- 3
clipping.mask <- raster(paste0(computer, "GIS_data/Perth-Peel/mask files/perth.clipping.mask.tif"))
  clipping.mask[!is.na(clipping.mask)] <- 3

```

The variables in the create.raster function are defined below:
s = input shapefile
mask.raster = the background raster that you are converting into
raster.label = label that is used as the filename, raster name and plot title
value = the value that cells are to be given where the shapefile is present in the landscape
transform = a logical value that determines whether the function transforms the shapefile data prior to conversion.  If set to TRUE, then you need to make sure that the CRS values are set correctly.

```{r create raster function}
create.raster <- function (s, mask.raster, raster.label, value, transform=TRUE, save=TRUE, plot=TRUE) {
    
   if(transform==TRUE) {
     proj4string(s) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
    s <- spTransform(s, CRS("+proj=utm +zone=50 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))
     }
    
  r <- rasterize(s,mask.raster)
    r[!is.na(r)] <- value
    r <- mask(merge(r,mask.raster),mask.raster)
    
    names(r) <- rep(raster.label,length(names(r)))
    
    if(save==TRUE) writeRaster(r,filename=paste0(output.directory,raster.label), format="GTiff", overwrite=T)
    
    if(plot==TRUE) plot(r, main=raster.label, zlim=c(1,6))

    return(r)
  }

```

```{r Class of Action}
  coa.urban <- readShapePoly(paste0(envi.directory,"CoA/urban_coa_20150204/urban_coa_20150204.shp"))
  coa.urban.raster <- create.raster(coa.urban,perth, raster.label="scenario1_urban", value=1, transform=FALSE)
  
  coa.industrial <- readShapePoly(paste0(envi.directory,"CoA/DoP_industrial_coa_20150204/DoP_industrial_coa_20150204.shp"))
  coa.industrial.raster <- create.raster(coa.industrial,perth, raster.label="scenario2_industrial", value=1, transform=FALSE)
  
  coa.ruralres <- readShapePoly(paste0(envi.directory,"CoA/DoP_rural_res_coa_20150130/DoP_rural_res_coa_20150130.shp"))
  coa.ruralres.raster <- create.raster(coa.ruralres,perth, raster.label="scenario3_ruralres", value=1, transform=FALSE)
  
  coa.infrastructure <- readShapePoly(paste0(envi.directory,"CoA/DoP_Infrastructure_Footprint_20151019/DoP_Infrastructure_Footprint_20151019.shp"))
  coa.infrastructure.raster <- create.raster(coa.infrastructure,perth, raster.label="scenario4_infrastructure", value=1, transform=FALSE)
  
  coa.pines <- readShapePoly(paste0(envi.directory,"CoA/DoP_Pines_COA_20151224/DoP_Pines_COA_20151224.shp"))
  coa.pines.raster <- create.raster(coa.pines,perth, raster.label="scenario5_pines", value=1, transform=TRUE)
  
  coa.brm <- readShapePoly(paste0(envi.directory,"CoA/DMP_BRM_Class_of_Action_SA_Public_20151209/DMP_BRM_Class_of_Action_SA_Public_20151209.shp"))
  coa.brm.raster <- create.raster(coa.brm[coa.brm$ColourCode!="Red",],perth, raster.label="scenario6_brm", value=1, transform=TRUE)
  
```

```{r cumulative scenario}
  
  coa.cumulative.raster <- calc(stack(coa.urban.raster,coa.industrial.raster,coa.ruralres.raster,coa.infrastructure.raster,coa.pines.raster,coa.brm.raster),min,filename=paste0(output.directory,"scenario7_cumulative"),format="GTiff",overwrite=T)
  plot(coa.cumulative.raster, main="coa.cumulative.raster", zlim=c(1,5))

```

```{r urban iterations}
# exclude industrial & rural residential from these to make fair comparison with coa_urban
  
  #EIA1
  dop.eia1 <- readShapePoly(paste0(envi.directory,"DOP/DoP_EIA1_20130809.shp"))
  dop.eia1.raster <- create.raster(dop.eia1[dop.eia1$Land_Type!="Undeveloped Urban" & dop.eia1$Land_Type!="Existing Industrial" & dop.eia1$Land_Type!="Existing Industrial" & dop.eia1$Land_Type!="Potential Industrial",],perth, raster.label="scenario1_1_eia1", value=1, transform=FALSE)
 
  # EIA2 
  dop.eia2 <- readShapePoly(paste0(envi.directory,"DOP/DoP_EIA2_20130809.shp"))
  dop.eia2.raster <- create.raster(dop.eia2[dop.eia2$Land_Type!="Developed Urban" & dop.eia2$Land_Type!="Existing Industrial" & dop.eia2$Land_Type!="Potential Industrial",], perth, raster.label="scenario1_2_eia2", value=1, transform=FALSE)
  
  # Option 1 without Keralup
  dop.EIA3_Option1_30102013 <- readShapePoly(paste0(envi.directory,"DOP/DOP_EIA3_Option1_30102013.shp"))
  dop.EIA3_Option1_30102013.raster <- create.raster(dop.EIA3_Option1_30102013, perth, raster.label="scenario1_3a_eia3_option1",value=1,transform=FALSE)
  

  # Option 2 with Keralup
  dop.EIA3_Option2_30102013 <- readShapePoly(paste0(envi.directory,"DOP/DOP_EIA3_Option2_30102013.shp"))
  dop.EIA3_Option2_30102013.raster <- create.raster(dop.EIA3_Option2_30102013, perth, raster.label="scenario1_3b_eia3_option2",value=1,transform=FALSE)
  
  # current development footprint
  dop.development.footprint <- readShapePoly(paste0(envi.directory,"DOP/DoP_Development_Footprint_20140310.shp"))
  dop.development.footprint.raster <- create.raster(dop.development.footprint[dop.development.footprint$DEVELOPED=="Urban Undeveloped",], perth, raster.label="scenario1_4_development_footprint",value=1,transform=FALSE)
  
```

```{r use old iteration scenarios}

  eia1.raster <- raster(paste0(computer,"GIS_data/Perth-Peel/zonation/100m/scenarios/final/CL_EIA1.tif"))
  eia2.raster <- raster(paste0(computer,"GIS_data/Perth-Peel/zonation/100m/scenarios/final/CL_EIA2.tif"))
  eia3a.raster <- raster(paste0(computer,"GIS_data/Perth-Peel/zonation/100m/scenarios/final/CL_EIA3O1.tif"))
  eia3b.raster <- raster(paste0(computer,"GIS_data/Perth-Peel/zonation/100m/scenarios/final/CL_EIA3O2.tif"))
  eia4.raster <- raster(paste0(computer,"GIS_data/Perth-Peel/zonation/100m/scenarios/final/scenario3b_urban.tif"))
    names(eia4.raster) <- "CL_EIA4"
  
  eia.rasters <- stack(eia1.raster,eia2.raster,eia3a.raster,eia3b.raster,eia4.raster)
  
  for(i in 1:nlayers(eia.rasters)){
    r <- eia.rasters[[i]]
    writeRaster(r, filename=paste0(computer, "GIS_data/Perth-Peel/zonation/100m/Conservation_Letters/final/scenarios/", names(r)), format="GTiff", overwrite=T)
    r <- mask(r,clipping.mask,filename=paste0(computer, "GIS_data/Perth-Peel/zonation/100m/Conservation_Letters/final/scenarios/", names(r),"_clipped"), format="GTiff", overwrite=T) 
    
  }
  
```


```{r merge eia scenarios with existing urban}
  # existing urban
  dop.existing.urban.raster <- create.raster(dop.eia1[dop.eia1$Land_Type=="Developed Urban",],perth, raster.label="existing.urban", value=1, transform=FALSE, save=FALSE)
    
  # coa.urban (without existing)
  coa.urban.no.existing <- sum(coa.urban.raster,dop.existing.urban.raster)
    coa.urban.no.existing[coa.urban.no.existing==6] <- 3
    coa.urban.no.existing[coa.urban.no.existing==2] <- 3
    coa.urban.no.existing[coa.urban.no.existing==4] <- 1
    names(coa.urban.no.existing) <- "coa.urban.no.existing"
    writeRaster(coa.urban.no.existing,filename=names(coa.urban.no.existing),format="GTiff", overwrite=T)
  
  dop.existing.urban.raster[dop.existing.urban.raster==3] <- NA
  
  dop.eia1.existing.raster <- merge(dop.existing.urban.raster,dop.eia1.raster,filename="scenario1_1_eia1_existing", format="GTiff", overwrite=T)
  
  dop.eia2.existing.raster <- merge(dop.existing.urban.raster,dop.eia2.raster,filename="scenario1_2_eia2_existing", format="GTiff", overwrite=T)
  
  dop.EIA3_Option1_30102013.existing.raster <- merge(dop.existing.urban.raster,dop.EIA3_Option1_30102013.raster,filename="scenario1_3a_eia3_option1_existing", format="GTiff", overwrite=T)
  
  dop.EIA3_Option2_30102013.existing.raster <- merge(dop.existing.urban.raster,dop.EIA3_Option2_30102013.raster,filename="scenario1_3b_eia3_option2_existing", format="GTiff", overwrite=T)
  
 dop.development.footprint.existing.raster <- merge(dop.existing.urban.raster,dop.development.footprint.raster,filename="scenario1_4_development_footprint_existing", format="GTiff", overwrite=T) 
  
```

```{r directions2031-GGP comparison}
# Green Growth Plan
 master.plan <- readShapePoly(paste0(envi.directory,"DoP/DoP_MasterPlan_UrbanExpansionSites_20160314.shp"))
    master.plan.raster <- create.raster(master.plan, perth, raster.label = "GGP_UrbanExpansion", value=1,transform=FALSE)
 
# retention zones
  specific.commitments <- readOGR(paste0(envi.directory,"DoP"),"DoP_SpecificCommitmentsArea_20160324")
    specific.commitments.raster <- create.raster(specific.commitments, perth, raster.label = "SpecificCommitmentsArea", value=1,transform=FALSE,save=FALSE)
    specific.commitments.raster[specific.commitments.raster==3] <-NA
    
    # remove retention zones from master planning scenario
    master.plan.raster[specific.commitments.raster==1] <- 3
    writeRaster(master.plan.raster,filename="GGP_UrbanExpansion_NoRetentionZones", format="GTiff", overwrite=T)
    
 
 # Directions 2031
 urban.expansion <- readShapePoly(paste0(envi.directory,"DoP/DoP_UrbanExpansionInvestigation_20130328.shp"))
  urban.expansion.raster <- create.raster(urban.expansion, perth, raster.label = "D2031_UrbanExpansion", value=1,transform=FALSE)
 
 
```


Make iterative cumulative scenarios

```{r iterative cumulative scenarios}
 
 old.scenarios <- dir(paste0(envi.directory,"zonation/100m/scenarios/final/"),pattern="tif$",full.names=T)
 
 scenario.data <- data.frame(cum.scenario=paste0("scenario_cumulative",seq(1:3)),urban.industrial=NA,infrastructure=NA,brm=NA,pines=NA)
  scenario.data[1,2:5] <- paste0(c("scenario1_eia1", "scenario8_infrastructure", "scenario4_brm.eia1","scenario7_pines"),".tif")
  scenario.data[2,2:5] <- paste0(c("scenario2_eia3", "scenario8_infrastructure", "scenario5_brm.eil4","scenario7_pines"),".tif")
  scenario.data[3,2:5] <- paste0(c("scenario3_development.footprint.lgz", "scenario8_infrastructure", "scenario6_brm.extraction.areas","scenario7_pines"),".tif")
  
  for(j in 1:nrow(scenario.data)){
    name <- as.character(scenario.data$cum.scenario[j])
    u <- raster(old.scenarios[grepl(scenario.data[j,"urban.industrial"],old.scenarios)])
    i <- raster(old.scenarios[grepl(scenario.data[j,"infrastructure"],old.scenarios)])
    b <- raster(old.scenarios[grepl(scenario.data[j,"brm"],old.scenarios)])
    p <- raster(old.scenarios[grepl(scenario.data[j,"pines"],old.scenarios)])
    
    cum <- min(u,i,b,p)
      cum[cum<3] <- 1
      names(cum) <- name
    assign(name, cum)
    writeRaster(cum,paste0(output.directory,name,".tif"),overwrite=T)
    plot(cum,main=name)
    
    rm(u,i,b,p,cum)
  }

```

```{r Existing protected areas}
    
# Existing IUCN 1-4a
  iucn1_4a <- readShapePoly(paste0(envi.directory, "CoA/Conservation_Package_Release_Web_March2016/DPaW_ExistingConsRes_IUCN1a_4_20150630.shp"))
    iucn1_4a.raster <- create.raster(iucn1_4a, perth, raster.label = "DPaW_ExistingConsRes_IUCN1a_4_20150630", value=6,transform=T,save=F)

  # DPAW Managed Lands
  managed.lands <- readShapePoly(paste0(envi.directory, "DEC/DEC_ManagedLandsAndWaters_20120601.shp"))
    managed.lands$IUCN[managed.lands$CATEGORY=="State Forest"] <- 6
    
    managed.lands4 <- create.raster(managed.lands[is.na(managed.lands$IUCN),],perth, raster.label="DEC_ManagedLandsAndWaters_20120601_No IUCN status", value=4,save=F)
    
    managed.lands <- managed.lands[!is.na(managed.lands$IUCN),]
    
    # State Forests (IUCN 6)
    managed.lands5 <- create.raster(managed.lands[managed.lands$IUCN=="6",],perth, raster.label="DEC_ManagedLandsAndWaters_20120601_IUCN 6", value=5,save=F)
    
# Heritage
world.heritage <- readShapePoly(paste0(envi.directory, "SEWPaC/PP_MNES_extract/PP_WHA_clip.shp"))
  world.heritage6 <- create.raster(world.heritage,perth, raster.label="PP_WHA_clip", value=6,save=F)

national.heritage <- readShapePoly(paste0(envi.directory, "SEWPaC/PP_MNES_extract/PP_NHL_clip.shp"))
  national.heritage6 <- create.raster(national.heritage,perth, raster.label="PP_NHL_clip", value=6,save=F)

commonwealth.heritage <- readShapePoly(paste0(envi.directory, "SEWPaC/PP_MNES_extract/PP_CHL_clip.shp"))
  commonwealth.heritage6 <- create.raster(commonwealth.heritage,perth, raster.label="PP_CHL_clip", value=6,save=F)

# Bush Forever
bushforever.SP28 <- readShapePoly(paste0(envi.directory,"DOP/DoP_StatePlanningPolicy28_BushForeverReserves_20120823.shp"))
  bushforever4 <- create.raster(bushforever.SP28,perth, raster.label="DoP_StatePlanningPolicy28_BushForeverReserves_20120823", value=4, transform=FALSE,save=F)

#MRS
parks.rec <- readShapePoly(paste0(envi.directory,"MRS_PRS/mrspoly_20120928.shp"))
  parks.rec4 <- create.raster(parks.rec[(parks.rec$DESCRIPTIO=="PARKS & RECREATION" | parks.rec$DESCRIPTIO=="PARKS & RECREATION RESTRICTED"), ],perth, raster.label="mrspoly_20120928", value=4, transform=FALSE,save=F)

#PRS
open.space <- readShapePoly(paste0(envi.directory,"MRS_PRS/prspoly_20120928.shp"))
  open.space4 <- create.raster(open.space[open.space$DESCRIPTIO=="REGIONAL OPEN SPACE" , ],perth, raster.label="prspoly_20120928", value=4, transform=FALSE,save=F)

protected.areas <- calc(stack(iucn1_4a.raster,managed.lands4,managed.lands5,world.heritage6,national.heritage6,commonwealth.heritage6,bushforever4,parks.rec4,open.space4),fun=max, filename=paste0(envi.directory,"protected_areas"), format="GTiff",overwrite=T)
  plot(protected.areas, main="Protected areas", zlim=c(3,6))

```

1. Convert Phase 1 & Phase 2 to raster, clipping to the PPSA boundary
2a. Flatten Existing PA layer to one value (6)
2b. Extract high security protected areas (6)
3a. Merge CR with PA - PA (6), Phase 1 (5), Phase 2 (4), rest (3), with CR status overwriting PA status.
3b. Merge CR with PA-high - PA-high (6), Phase 1 (5), Phase 2 (4), rest (3), with CR status overwriting PA status.

```{r proposed conservation reserves}
  
  conservation.reserves <-  readShapePoly( paste0(envi.directory, "CoA/DPaw_proposed_conservation_reserves/DPaW_Proposed_GGP_Conservation_Reserves_20160316.shp"))
  
  CR.phase1.raster <- create.raster(conservation.reserves[conservation.reserves$PRPhase=="Phase 1",], perth,  raster.label="CR_phase1", value=5, transform=TRUE,save = FALSE)
    CR.phase1.raster[CR.phase1.raster==3] <- NA
  CR.phase2.raster <- create.raster(conservation.reserves[conservation.reserves$PRPhase=="Phase 2",], perth, raster.label="CR_phase2", value=4, transform=TRUE,save=FALSE)
    CR.phase2.raster[CR.phase2.raster==3] <- NA
  
  # load existing PA raster layer
  PA.raster <- raster(paste0("protected_areas.tif"))
    
  # flatten all PA security levels
  PA.raster.flat <- PA.raster
    PA.raster.flat[PA.raster.flat>3] <- 6
    
  PA.raster.high <- PA.raster
    PA.raster.high[PA.raster.high!=6] <- 3
  writeRaster(PA.raster.high,filename="high_protected_areas", format="GTiff", overwrite=T)
    PA.raster.high[PA.raster.high!=6] <- NA
    
  CR.raster <- merge(CR.phase1.raster,CR.phase2.raster,perth,filename="conservation_reserves", format="GTiff", overwrite=T)
  # CR.PA.raster <- merge(CR.phase1.raster,CR.phase2.raster,PA.raster.flat,filename="protected_areas_conservation_reserves", format="GTiff", overwrite=T)
  CR.PA.high.raster <- merge(PA.raster.high,CR.phase1.raster,CR.phase2.raster,perth,filename="high_protected_areas_conservation_reserves", format="GTiff", overwrite=T)

  conservation.rasters <- c(PA.raster,CR.raster,CR.PA.high.raster)
  
  par(mfrow=c(1,3))
  for(i in seq(conservation.rasters)) {
    plot(conservation.rasters[[i]], main=names(conservation.rasters[[i]]), zlim=c(3,6),legend=F,axes=F,box=F)
  }
  
  
```



All the scenarios need to be clipped by the vegetation layer to remove the areas that are already developed and, thus, not being considered.


```{r clip scenarios by clipping mask}
  scenarios <- dir(output.directory,pattern=".tif$")
    scenarios <- scenarios[!grepl("clipped",scenarios)]
  
  for (s in seq(scenarios)){
    r <- raster(scenarios[s])
    r <- mask(r,clipping.mask,filename=paste0(computer, "GIS_data/Perth-Peel/zonation/100m/Conservation_Letters/final/scenarios/", names(r),"_clipped"), format="GTiff", overwrite=T)
    plot(r, main=names(r),zlim=c(1,6))
  }

```



*This file was last updated on 12 January 2016 and last run on `r format(Sys.time(), "%d %B %Y")`.*